Unsupervised embeddings of organizations capture the latent structure of scientific mobility

### 2020-06-10 paper/current/main.tex
effective distance -> functional distance

%
% Subsection: Embedding performance
%
\subsection*{Embeddings estimate the effective distance between organizations}

The effective distance estimated by our embedding model explains and predicts scientific mobility best.
The amount of variance explained by embedding distance ($R^{2} = 0.46$, Fig.~\ref{fig:gravity_performance}a) is more than twice that of geographic distance ($R^{2} = 0.22$, Fig.~\ref{fig:gravity_performance}b).
This pattern holds for the variance of only domestic (Fig.~\ref{fig:supp:gravity_correlation}c) and international mobility flows (Fig.~\ref{fig:supp:gravity_correlation}d).
After fitting model parameters from the data, the embedding distance produces better predictions of expected flux than does geographic distance globally (Fig.~\ref{fig:gravity_performance}c--d) and for domestic (Fig.~\ref{fig:supp:gravity_correlation}e--f) and international mobility(Fig.~\ref{fig:supp:gravity_correlation}g--h).
We also examine gravity model with dot product in the embedding space, the amount of variance explain ($R^{2} = 0.34$) by dot product outperforms that of geographic distance (Fig.~\ref{fig:supp:gravity_dot}).
Similarly, embedding distance outperforms ``diffusion''  distances such as those derived from calculating personalized-page rank scores (see supporting materials) on the underlying mobility network (Fig.~\ref{fig:supp:gravity_pprcos}, Fig.~\ref{fig:supp:gravity_pprjsd}).
Geographic distance alone is a poor perspective for understanding international scientific mobility---rather, by learning a distance directly from the data, it is possible to better explain, and better predict flux between organizations.

@@

Given that embeddings capture effective distance between organizations, we can now examine the global structure of scientific mobility.
Because the embeddings are defined in a high-dimensional space, we apply a widely-used dimensionality reduction technique (UMAP~\autocite{mcinnes2018umap}) to visualize the overall structure (Fig.~\ref{fig:projection}a).
Globally, the effect of geography is conspicuous; organizations tend to form clusters based on their national affiliations and national clusters tend to be near their geographic neighbors. 
At the same time, the embedding space also reflects a mix of geographic, historic, cultural, and linguistic relationships between regions much more clearly than alternative network visualizations (Fig.~\ref{fig:supp:network_vis}).

>>>>

%
% Subsection: Embedding performance
%
\subsection*{Embeddings provide a functional distance between organizations}

The functional distance derived from the  embedding model explains and predicts scientific mobility best.
The amount of variance explained by embedding distance ($R^{2} = 0.46$, Fig.~\ref{fig:gravity_performance}a) is more than twice that of geographic distance ($R^{2} = 0.22$, Fig.~\ref{fig:gravity_performance}b).
This pattern holds for the variance of only domestic (Fig.~\ref{fig:supp:gravity_correlation}c) and international mobility flows (Fig.~\ref{fig:supp:gravity_correlation}d).
After fitting model parameters from the data, the embedding distance produces better predictions of expected flux than does geographic distance globally (Fig.~\ref{fig:gravity_performance}c--d) and for domestic (Fig.~\ref{fig:supp:gravity_correlation}e--f) and international mobility(Fig.~\ref{fig:supp:gravity_correlation}g--h).
We also examine gravity model with dot product in the embedding space, the amount of variance explain ($R^{2} = 0.34$) by dot product outperforms that of geographic distance (Fig.~\ref{fig:supp:gravity_dot}).
Similarly, embedding distance outperforms ``diffusion''  distances such as those derived from calculating personalized-page rank scores (see supporting materials) on the underlying mobility network (Fig.~\ref{fig:supp:gravity_pprcos}, Fig.~\ref{fig:supp:gravity_pprjsd}).
Geographic distance alone is a poor perspective for understanding international scientific mobility---rather, by learning a distance directly from the data, it is possible to better explain, and better predict flux between organizations.

@@

Given that embeddings capture a functional distance between organizations, we can now examine the global structure of scientific mobility.
Because the embeddings are defined in a high-dimensional space, we apply a widely-used dimensionality reduction technique (UMAP~\autocite{mcinnes2018umap}) to visualize the overall structure (Fig.~\ref{fig:projection}a).
Globally, the effect of geography is conspicuous; organizations tend to form clusters based on their national affiliations and national clusters tend to be near their geographic neighbors. 
At the same time, the embedding space also reflects a mix of geographic, historic, cultural, and linguistic relationships between regions much more clearly than alternative network visualizations (Fig.~\ref{fig:supp:network_vis}).
